---
title: "DOM indices"
author: "Mackenzie Fiss"
date: "1/27/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cmocean)
library(ggplot2)
library(GGally)
```

*Formatting*

```{r}
all.chem <- FTIC.data.noSDL
mz <- read.csv("mass.csv", header = T)
all.chem <- join(all.chem, mz, by = "Formula")
rownames(all.chem) <- all.chem$Formula

sample.names <- data.frame(Sample.ID = rownames(norm.FTIC))

all.dom <- data.frame(t(norm.FTIC))
```

# Ideg - Flerus et al 2012

```{r}
d.neg <- c("C21H26O11", "C17H20O9", "C19H22O10", "C20H22O10", "C20H24O11")
d.pos <- c("C13H18O7", "C14H20O7", "C15H22O7", "C15H22O8", "C16H24O8")
ftdom.d.neg <- (all.dom)[rownames(all.dom)%in%d.neg,]
d.neg.sums <- colSums(ftdom.d.neg)
ftdom.d.pos <-  (all.dom)[rownames(all.dom)%in%d.pos,]
d.pos.sums <- colSums(ftdom.d.pos)
ideg <- as.data.frame(d.neg.sums/(d.neg.sums+d.pos.sums))
ideg <- cbind(ideg, rownames(ideg))
colnames(ideg) <- c("ideg", "names")

ideg$Date <- PCA.scores.noSDL$Date
```

## Ideg over time

```{r}
# Ideg over time
ggplot(ideg) + 
  geom_point(aes(x = Date, y = ideg, color = ideg), shape = 20, alpha = 0.7, size = 4) + 
  labs(x = "Month (2013 - 2015)", y = bquote(I[deg]*" (Flerus et al 2012)"), color = bquote(I[deg])) + 
  scale_y_continuous(limits = c(0.3, 0.55))+
  scale_color_cmocean(name = "ice", direction = -1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black", frame.linewidth = 0.25), limits = c(0.3, 0.6), breaks = seq(0.3, 0.6, 0.1)) + 
  #scale_x_date(date_labels = "%b", date_breaks = "8 weeks", limits = as.Date(c("2013-08-05","2015-02-26"))) + 
  theme_linedraw() +  
  theme(panel.grid.major = element_line(color = "#EBEBEB"), panel.grid.minor = element_blank())
```

## Ideg of Gonzalo's porewater

```{r}
porewater.dom <- data.frame(t(norm.porewater.FTIC))

ftdom.d.neg.pore <- (porewater.dom)[rownames(porewater.dom)%in%d.neg,]
d.neg.sums.pore <- colSums(ftdom.d.neg.pore)
ftdom.d.pos.pore <-  (porewater.dom)[rownames(porewater.dom)%in%d.pos,]
(d.pos.sums.pore <- colSums(ftdom.d.pos.pore))
ideg.pore <- as.data.frame(d.neg.sums.pore/(d.neg.sums.pore+d.pos.sums.pore))

#ideg.pore <- cbind(ideg, rownames(ideg))
#colnames(ideg) <- c("ideg", "names")
```


# %IOS - Island of stability - Lechtenfeld et al 2014

```{r}
ios.chem2 <- subset(all.chem, all.chem$N==0 & all.chem$S==0 & all.chem$P==0)
ios.cho.dom <- as.data.frame((all.dom)[rownames(all.dom)%in%rownames(ios.chem2),])
ios.chem <- subset(all.chem, all.chem$HC>1.04 & all.chem$HC<1.3)
ios.chem <- subset(ios.chem, ios.chem$OC>0.42 & ios.chem$OC<0.62)
ios.chem <- subset(ios.chem, ios.chem$mz>332 & ios.chem$mz<388 | ios.chem$mz>446 & ios.chem$mz<548)
ios.dom <- as.data.frame((all.dom)[rownames(all.dom)%in%rownames(ios.chem),])
ios.percent <- as.data.frame(colSums(ios.dom)/colSums(ios.cho.dom)*100)
colnames(ios.percent) <- "IoS.percent"
```

# Iterr - Medeiros et al 2016

```{r}
t.neg <- c("C13H12O5","C15H14O4","C14H12O5","C14H14O5","C13H12O6","C16H15O4", "C15H14O5","C14H12O6", "C15H16O5","C14H14O6","C16H14O5","C16H16O5","C15H14O6","C15H16O6","C14H14O7","C17H16O5","C16H14O6", "C17H18O5", "C16H16O6","C15H14O7","C17H16O6","C16H14O7","C18H18O6","C17H16O7","C17H18O7", "C18H16O7", "C18H18O7","C17H16O8","C19H18O7","C20H20O7","C19H18O8","C20H18O9","C19H16O10","C21H20O9","C20H18O10","C22H22O9","C21H20O10","C23H22O10","C24H24O10","C25H26O10")
t.pos <- c("C15H19NO6", "C15H21NO6", "C17H21NO7", "C17H23NO7", "C17H22O8", "C16H21NO8", "C17H20N2O7", "C17H19NO8", "C18H23NO7", "C17H21NO8", "C18H24O8", "C16H19NO9", "C17H23NO8","C17H22O9", "C17H24O9", "C18H21NO8","C17H19NO9", "C18H23NO8", "C18H22O9", "C17H21NO9","C18H24O9", "C18H20N2O8", "C18H21NO9", "C19H24O9", "C18H23NO9", "C18H22O10", "C18H24O10", "C20H24O9", "C19H22O10", "C20H26O9", "C19H24O10", "C19H26O10","C20H24O10","C20H26O10", "C19H24O11", "C20H24O11", "C20H26O11", "C20H26O12", "C22H28O11", "C21H28O12")
ftdom.t.neg <- (all.dom)[rownames(all.dom)%in%t.neg,]
t.neg.sums <- colSums(ftdom.t.neg)
ftdom.t.pos <- (all.dom)[rownames(all.dom)%in%t.pos,]
t.pos.sums <- colSums(ftdom.t.pos)
Iterr <- data.frame(Iterr = t.neg.sums/(t.neg.sums + t.pos.sums))
```

# Molecular lability boundary (MLB) from Andrilli et al 2023

## Overall MLB

```{r}
# % MLB
high.HC <- subset(FTIC.data.noSDL, HC >= 1.5, select = c("Formula", "comp2"))

high.HC.samples <- subset(norm_intensity_matrix, subset = (rownames(norm_intensity_matrix) %in% rownames(high.HC)))

overall.MLB <- data.frame(perc.MLB = (colSums(high.HC.samples > 0)/4431)*100)
```

### Overall MLB relationships

```{r}
overall.MLB <- cbind(PCA.scores.noSDL[c(1:31,38,39)], RDA.scores.noSDL[38:40], overall.MLB)

# MLB and degradation index over time
ggplot(overall.MLB) + 
  geom_point(aes(x = Datetime, y = ideg, color = perc.MLB), shape = 20, alpha = 0.7, size = 4) + 
  labs(x = "Month (2013 - 2015)", 
       y = bquote(I[DEG]), 
       color = bquote("%"*MLB[L]),
       caption = bquote("%"*MLB[L]*"= molecular lability bounday from D'Andrilli et al 2023, "*I[DEG]*"= degradation index from Flerus et al 2012")) + 
  scale_y_continuous(limits = c(0.3, 0.55))+
  scale_color_cmocean(name = "ice", direction = -1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black", frame.linewidth = 0.25), limits = c(7, 12), breaks = seq(7, 12, 1)) + 
  #scale_x_date(date_labels = "%b", date_breaks = "8 weeks", limits = as.Date(c("2013-08-05","2015-02-26"))) + 
  theme_linedraw() +  
  theme(panel.grid.major = element_line(color = "#EBEBEB"), panel.grid.minor = element_blank())

ggplot(overall.MLB,
       aes(x = avg.OC, y = PC1, color = perc.MLB)) + 
  geom_point(shape = 20, alpha = 0.7, size = 4) + 
  labs(x = "Intensity-weighted average O/C", 
       y = "PC1 (25.9%)", 
       color = bquote("%"*MLB[L]), 
       caption = bquote("%"*MLB[L]*"= molecular lability bounday from D'Andrilli et al 2023")) + 
  scale_x_continuous(limits=c(0.33, 0.45), breaks=seq(0.33, 0.45, 0.03)) +
  scale_y_continuous(limits=c(-0.2, 0.2)) +
  scale_color_cmocean(name = "matter", direction=1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black", frame.linewidth = 0.25), limits = c(7,12)) + 
  theme_linedraw() + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  stat_cor(method = "pearson",cor.coef.name = c("R"), color = "black")

ggsave("PC1_OC_MLB.pdf", width = 6, height = 4)

ggpairs(MLB,
        columns = c(37,34:36,32:33),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("MLB", "RDA1", "RDA2", "RDA3", "PC1", "PC2")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))

```

## Per heteroatom group MLB

```{r}
FTIC.data.noSDL$comp2 <- character(length(FTIC.data.noSDL$C))

for (i in 1:length(FTIC.data.noSDL$C)) {
          ifelse(FTIC.data.noSDL$C[i] > 0 & FTIC.data.noSDL$H[i] > 0 & FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHO",
          ifelse(FTIC.data.noSDL$C[i] > 0 & FTIC.data.noSDL$H[i] > 0 & FTIC.data.noSDL$O[i] == 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CH", 
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 1 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHON1",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 2 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHON2",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 3 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHON3",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 4 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHON4",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 1, FTIC.data.noSDL$comp2[i] <- "CHOP1", 
          ifelse(FTIC.data.noSDL$O[i] == 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 1, FTIC.data.noSDL$comp2[i] <- "CHP1", 
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 1 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHOS1", 
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 2 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHOS2",
          ifelse(FTIC.data.noSDL$O[i] == 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 1 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHS1",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 1 & FTIC.data.noSDL$S[i] == 0 & FTIC.data.noSDL$P[i] == 1, FTIC.data.noSDL$comp2[i] <- "CHON1P1", 
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 1 & FTIC.data.noSDL$S[i] == 1 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHON1S1",
          ifelse(FTIC.data.noSDL$O[i] > 0 & FTIC.data.noSDL$N[i] == 0 & FTIC.data.noSDL$S[i] == 1 & FTIC.data.noSDL$P[i] == 1, FTIC.data.noSDL$comp2[i] <- "CHOP1S1",
          ifelse(FTIC.data.noSDL$O[i] == 0 & FTIC.data.noSDL$N[i] == 1 & FTIC.data.noSDL$S[i] == 1 & FTIC.data.noSDL$P[i] == 0, FTIC.data.noSDL$comp2[i] <- "CHN1S1",
          FTIC.data.noSDL$comp2[i] <- "oops")))))))))))))))
}

high.HC.samples.comp <- cbind(high.HC.samples, high.HC[2])

table(FTIC.data.noSDL$comp2)
table(high.HC.samples.comp$comp2)

# CHO
CHO.samples <- subset(high.HC.samples.comp, comp2 == "CHO")
CHO.MLB <- data.frame(CHO.MLB = (colSums(CHO.samples[,1:322] > 0)/(sum(FTIC.data.noSDL$comp2=="CHO")))*100)

# CHON
CHON1.samples <- subset(high.HC.samples.comp, comp2 == "CHON1")
CHON1.MLB <- data.frame(CHON1.MLB = (colSums(CHON1.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON1"))*100)
CHON2.samples <- subset(high.HC.samples.comp, comp2 == "CHON2")
CHON2.MLB <- data.frame(CHON2.MLB = (colSums(CHON2.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON2"))*100)
CHON4.samples <- subset(high.HC.samples.comp, comp2 == "CHON4")
CHON4.MLB <- data.frame(CHON4.MLB = (colSums(CHON4.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON4"))*100)

CHON.samples <- subset(high.HC.samples.comp, comp2 == "CHON1" | comp2 == "CHON2" | comp2 == "CHON4")
CHON.MLB <- data.frame(CHON.MLB = (colSums(CHON.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON1" | FTIC.data.noSDL$comp2=="CHON2" | FTIC.data.noSDL$comp2=="CHON4"))*100)

# CHOP
CHOP1.samples <- subset(high.HC.samples.comp, comp2 == "CHOP1")
CHOP1.MLB <- data.frame(CHOP1.MLB = (colSums(CHOP1.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHOP1"))*100)

# CHOS
CHOS1.samples <- subset(high.HC.samples.comp, comp2 == "CHOS1")
CHOS1.MLB <- data.frame(CHOS1.MLB = (colSums(CHOS1.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHOS1"))*100)
CHOS2.samples <- subset(high.HC.samples.comp, comp2 == "CHOS2")
CHOS2.MLB <- data.frame(CHOS2.MLB = (colSums(CHOS2.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHOS2"))*100)

CHOS.samples <- subset(high.HC.samples.comp, comp2 == "CHOS1" | comp2 == "CHOS2")
CHOS.MLB <- data.frame(CHOS.MLB = (colSums(CHOS.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHOS1" | FTIC.data.noSDL$comp2=="CHOS2"))*100)

# CHON1P1
CHON1P1.samples <- subset(high.HC.samples.comp, comp2 == "CHON1P1")
CHON1P1.MLB <- data.frame(CHON1P1.MLB = (colSums(CHON1P1.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON1P1"))*100)

# CHON1S1 
CHON1S1.samples <- subset(high.HC.samples.comp, comp2 == "CHON1S1")
CHON1S1.MLB <- data.frame(CHON1S1.MLB = (colSums(CHON1S1.samples[,1:322] > 0)/sum(FTIC.data.noSDL$comp2=="CHON1S1"))*100) 

hetero.MLB <- cbind(CHO.MLB,
                         CHON1.MLB,
                         CHON2.MLB,
                         CHON4.MLB,
                         CHON.MLB,
                         CHOP1.MLB,
                         CHOS1.MLB,
                         CHOS2.MLB,
                         CHOS.MLB)
MLB <- cbind(overall.MLB, hetero.MLB, RDA.scores.noSDL[c(32,34)])
colnames(MLB)[37] <- "overall.MLB"

MLB %>% 
  ggplot()+
  geom_point(aes(x = Shannon, y = perc.MLB, color = Ant.Weekly.Total.Discharge.m3s))+
  facet_wrap(~Season)+
  stat_cor(aes(x = Shannon, y = perc.MLB))

ggplot(MLB, 
       aes(x = Datetime, y = `PO4 (ugP/L)`, color = avg.OC)) + 
  geom_point(size = 2.5, alpha = 0.7) + 
  #labs(x = "Date") + 
  scale_x_datetime(limits=c(as.POSIXct("2013-10-01"),
                            as.POSIXct("2015-02-28")),
                   breaks = (seq(as.POSIXct("2013-11-01"),
                                 as.POSIXct("2015-02-01"), "3 months")),
                   labels = scales::label_date("%b %Y")) +
  #scale_y_continuous(limits=c(-0.2, 0.2)) +
  scale_color_cmocean(name = "deep", direction=1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black", frame.linewidth = 0.25)) + # limits=c(0.33, 0.45), breaks=seq(0.33, 0.45, 0.03)
  theme_linedraw() + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  stat_cor(method = "pearson",cor.coef.name = c("R"), color = "black")#+
  #facet_grid(~Season)
```

```{r}
RDA.scores.noSDL <- cbind(RDA.scores.noSDL, subset(GC_DD_MF_all[8:19], rownames(GC_DD_MF_all) %in% rownames(RDA.scores.noSDL)))
```


### Heteroatom MLB relationships

```{r}
# MLB and RDA1-3
ggpairs(MLB,
        columns = c(37,38,42,43,46,34:36),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("overall MLB", "CHO MLB", "CHON MLB", "CHOP MLB", "CHOS MLB", "RDA1", "RDA2", "RDA3")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))

# MLB and PC1-2
ggpairs(MLB,
        columns = c(37,38,42,43,46,32:33),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("overall MLB", "CHO MLB", "CHON MLB", "CHOP MLB", "CHOS MLB", "PC1", "PC2")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))

# specific MLB heteroatom groups and RDA3 and PC1
# MLB and PC1-2
ggpairs(MLB,
        columns = c(39,40,41,43,36,32),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("CHON1 MLB", "CHON2 MLB", "CHON4 MLB", "CHOP MLB", "RDA3", "PC1")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))
```

# Imap - mangrove porewater index from Michael Seidel (unpublished

```{r}
m.pos<-c("C18H24O7S","C20H30O8S","C20H28O9S","C18H26O7S","C16H24O7S","C18H24O8S","C18H26O8S","C19H26O8S","C13H12O7S","C19H26O9S", "C17H24O8S","C12H10O6S","C18H28O7S","C16H26O7S","C19H28O7S","C16H20O9S","C14H14O8S","C19H28O8S","C16H22O8S","C15H14O8S","C17H22O8S","C14H10O6S","C16H14O6S","C11H10O5S","C13H10O6S","C15H12O7S","C13H12O6S","C22H32O8S","C15H16O9S","C19H24O9S","C15H18O8S","C18H22O9S","C13H12O5S","C11H12O6S","C21H32O8S","C14H12O5S","C14H14O7S","C13H16O8S","C13H14O7S","C13H18O7S")
ftdom.m.neg <- (all.dom)[rownames(all.dom)%in%t.pos,] # take t.pos (d13C has a negative sign and meant are the Island of stability peaks) 
m.neg.sums <- colSums(ftdom.m.neg)
ftdom.m.pos <- (all.dom)[rownames(all.dom)%in%m.pos,]
m.pos.sums <- colSums(ftdom.m.pos)
Imap <- data.frame(Imap = m.pos.sums/(m.pos.sums + m.neg.sums))
```

## testing if SUP formulas match the Imap formulas (what he published the index as)

```{r}
mbind<-function(...){
 Reduce( function(x,y){cbind(x,y[match(row.names(x),row.names(y)),])}, list(...) )
}
```

```{r}
sup.formulas <- data.frame(sup.form = c("C11H10O5S","C11H12O6S","C13H12O5S","C12H10O6S","C14H12O5S","C13H10O6S","C13H12O6S","C14H10O6S","C13H12O7S","C13H14O7S","C13H18O7S","C14H14O7S","C13H16O8S","C16H14O6S","C15H12O7S","C14H14O8S","C15H14O8S","C15H18O8S","C16H24O7S","C16H26O7S","C15H16O9S","C16H22O8S","C18H24O7S","C17H22O8S","C18H26O7S","C16H20O9S","C17H24O8S","C18H28O7S","C18H24O8S","C19H28O7S","C18H26O8S","C18H22O9S","C19H26O8S","C19H28O8S","C19H24O9S","C19H26O9S","C20H30O8S","C20H28O9S","C21H32O8S","C22H32O8S"), 
                           row.names = c("C11H10O5S","C11H12O6S","C13H12O5S","C12H10O6S","C14H12O5S","C13H10O6S","C13H12O6S","C14H10O6S","C13H12O7S","C13H14O7S","C13H18O7S","C14H14O7S","C13H16O8S","C16H14O6S","C15H12O7S","C14H14O8S","C15H14O8S","C15H18O8S","C16H24O7S","C16H26O7S","C15H16O9S","C16H22O8S","C18H24O7S","C17H22O8S","C18H26O7S","C16H20O9S","C17H24O8S","C18H28O7S","C18H24O8S","C19H28O7S","C18H26O8S","C18H22O9S","C19H26O8S","C19H28O8S","C19H24O9S","C19H26O9S","C20H30O8S","C20H28O9S","C21H32O8S","C22H32O8S"))

m.pos.df <- data.frame(m.pos.form = c("C18H24O7S","C20H30O8S","C20H28O9S","C18H26O7S","C16H24O7S","C18H24O8S","C18H26O8S","C19H26O8S","C13H12O7S","C19H26O9S", "C17H24O8S","C12H10O6S","C18H28O7S","C16H26O7S","C19H28O7S","C16H20O9S","C14H14O8S","C19H28O8S","C16H22O8S","C15H14O8S","C17H22O8S","C14H10O6S","C16H14O6S","C11H10O5S","C13H10O6S","C15H12O7S","C13H12O6S","C22H32O8S","C15H16O9S","C19H24O9S","C15H18O8S","C18H22O9S","C13H12O5S","C11H12O6S","C21H32O8S","C14H12O5S","C14H14O7S","C13H16O8S","C13H14O7S","C13H18O7S"),
                       row.names = c("C18H24O7S","C20H30O8S","C20H28O9S","C18H26O7S","C16H24O7S","C18H24O8S","C18H26O8S","C19H26O8S","C13H12O7S","C19H26O9S", "C17H24O8S","C12H10O6S","C18H28O7S","C16H26O7S","C19H28O7S","C16H20O9S","C14H14O8S","C19H28O8S","C16H22O8S","C15H14O8S","C17H22O8S","C14H10O6S","C16H14O6S","C11H10O5S","C13H10O6S","C15H12O7S","C13H12O6S","C22H32O8S","C15H16O9S","C19H24O9S","C15H18O8S","C18H22O9S","C13H12O5S","C11H12O6S","C21H32O8S","C14H12O5S","C14H14O7S","C13H16O8S","C13H14O7S","C13H18O7S"))


s.matches <- mbind(sup.formulas, m.pos.df) # 40 matches -- all the same formulas
```


```{r}
all.index.sums <- as.data.frame(cbind(ideg[c(1,3)], ios.percent, Iterr, Imap))
```

# Iphoto and Ibio

```{r}
iphoto.formulas<- c("C16H16O7", "C19H20O7", "C16H20N2O9", "C23H26O8", "C24H30O8")
ibio.formulas<- c("C13H18O5", "C13H16O6", "C13H19NO6", "C13H17NO7", "C18H28O7")
resistant.formulas <- c("C14H16O8", "C17H24O8", "C16H22O9", "C19H28O9", "C19H28O10")

ftdom.iphoto <- (all.dom)[rownames(all.dom) %in% iphoto.formulas,]
iphoto.sums <- colSums(ftdom.iphoto)

ftdom.resistant <- (all.dom)[rownames(all.dom) %in% resistant.formulas,]
resistant.sums <- colSums(ftdom.resistant)

ftdom.ibio <- (all.dom)[rownames(all.dom) %in% ibio.formulas,]
ibio.sums <- colSums(ftdom.ibio)

iphoto <- data.frame(Iphoto = iphoto.sums/resistant.sums)
ibio <- data.frame(Ibio = ibio.sums/resistant.sums)

RDA.scores.noSDL <- cbind(RDA.scores.noSDL, iphoto, ibio)
PCA.scores.noSDL <- cbind(PCA.scores.noSDL, iphoto, ibio)
  
ggpairs(RDA.scores.noSDL,
        columns = c(38,39,40,76,77),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("RDA1", "RDA2", "RDA3", "Iphoto", "Ibio")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))

ggpairs(PCA.scores.noSDL,
        columns = c(38,39,43,44),
        aes(color = Season,
            alpha = 0.5),
        columnLabels = c("PC1", "PC2", "Iphoto", "Ibio")) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5)) +
  scale_color_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160")) +
  scale_fill_manual(values=c("#225256", "#619FAD", "#EAA6B0", "#DB6160"))

ggplot(RDA.scores.noSDL,
       aes(x = avg.OC, y = PC1, color = perc.MLB)) + 
  geom_point(shape = 20, alpha = 0.7, size = 4) + 
  labs(x = "Intensity-weighted average O/C", 
       y = "PC1 (25.9%)", 
       color = bquote("%"*MLB[L]), 
       caption = bquote("%"*MLB[L]*"= molecular lability bounday from D'Andrilli et al 2023")) + 
  scale_x_continuous(limits=c(0.33, 0.45), breaks=seq(0.33, 0.45, 0.03)) +
  scale_y_continuous(limits=c(-0.2, 0.2)) +
  scale_color_cmocean(name = "matter", direction=1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black", frame.linewidth = 0.25), limits = c(7,12)) + 
  theme_linedraw() + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  stat_cor(method = "pearson",cor.coef.name = c("R"), color = "black")

```

